---
title: "20231202_MixedProbesFullRun"
output: html_notebook
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath("/mnt/DATA/LairdLab/rojas/RhMc_SpTr_ClarkLab/")) 

{ 
  library(Seurat)
  library(tidyverse)
  library(Matrix)
  library(scales)
  library(cowplot)
  library(RCurl)
  #library(Matrix.utils) #
  library(clustree)
  library(ComplexHeatmap) 
  library(readxl)
  #library(EnhancedVolcano) #
  library(BiocManager)
  #library(AnnotationHub) #
  library(SoupX)
  library(scCustomize)
  library(DoubletFinder)
  library(harmony)
  library(patchwork)
}
```

# Load in the data and get 
Had to load in the spatial data with a unique source of code because seurat v5. I then run the VlnPlot and other QC Metrics

Important to note here. The way that I made the Black and white image was by having 2 locations for the outs folder, ONE is 
`"/mnt/usb/rojas/spatialtranscriptomics/spaceranger_outs/RhSpTr-emb020-021/emb020/outs/"` and the other is 
`"/mnt/DATA/LairdLab/rojas/data/spaceranger_outs/RhSpTr-emb020-021/emb020/outs/"`
And to make the Black and white I renamed the image that is BnW to be `tissue_lowres_image`. Then kept the other as `tissue_lowres_image_color`


```{r}

source("/mnt/DATA/LairdLab/rojas/functions/Rfunctions.R")

data_dir <- "/mnt/newusb/rojas/spatialtranscriptomics/spaceranger_outs/RhMa_spacerangerOuts-MIXEDprobes/"
list.files(data_dir)

emb019 <- Load10X_Spatial(data.dir = "/mnt/newusb/rojas/spatialtranscriptomics/spaceranger_outs/RhMa_spacerangerOuts-MIXEDprobes/emb019/outs/")
emb020 <- Load10X_Spatial(data.dir = "/mnt/newusb/rojas/spatialtranscriptomics/spaceranger_outs/RhMa_spacerangerOuts-MIXEDprobes/emb020/outs/")
emb021 <- Load10X_Spatial(data.dir = "/mnt/newusb/rojas/spatialtranscriptomics/spaceranger_outs/RhMa_spacerangerOuts-MIXEDprobes/emb021/outs/")
emb022 <- Load10X_Spatial(data.dir = "/mnt/newusb/rojas/spatialtranscriptomics/spaceranger_outs/RhMa_spacerangerOuts-MIXEDprobes/emb022/outs/")
emb029 <- Load10X_Spatial(data.dir = "/mnt/newusb/rojas/spatialtranscriptomics/spaceranger_outs/RhMa_spacerangerOuts-MIXEDprobes/emb029/outs/")
emb031 <- Load10X_Spatial(data.dir = "/mnt/newusb/rojas/spatialtranscriptomics/spaceranger_outs/RhMa_spacerangerOuts-MIXEDprobes/emb031/outs/")
emb032 <- Load10X_Spatial(data.dir = "/mnt/newusb/rojas/spatialtranscriptomics/spaceranger_outs/RhMa_spacerangerOuts-MIXEDprobes/emb032/outs/")

# Create a list of Seurat objects
seurat_objects <- list(emb019, emb020, emb021, emb022, emb029, emb031, emb032) 

# Loop through each object and plot
for (seurat_obj in seurat_objects) {
    # Violin plot
    p1 <- VlnPlot(seurat_obj, features = "nCount_Spatial", pt.size = 0) + NoLegend()

    # Spatial feature plot
    p2 <- SpatialFeaturePlot(seurat_obj, features = "nCount_Spatial", pt.size.factor = 1) + theme(legend.position = "right")
    print(p1) | print(p2)
}

```
#Comparing the SCTransform and Seurat Log Normalization
I wanted to make sure that I did not need to run SCTransform vs. the normal Log Normalization in Seurat. So I ran this code from [Seurat's Vignette](https://satijalab.org/seurat/articles/spatial_vignette#data-preprocessing) 


```{r eval=FALSE, include=FALSE}
# rerun normalization to store sctransform residuals for all genes
emb020 <- SCTransform(emb020, assay = "Spatial", return.only.var.genes = FALSE)
# also run standard log normalization for comparison
emb020 <- NormalizeData(emb020, verbose = FALSE, assay = "Spatial")
# Computes the correlation of the log normalized data and sctransform residuals with the
# number of UMIs
emb020 <- GroupCorrelation(emb020, group.assay = "Spatial", assay = "Spatial", layer = "data", do.plot = FALSE)
emb020 <- GroupCorrelation(emb020, group.assay = "Spatial", assay = "SCT", slot = "scale.data", do.plot = FALSE)
p1 <- GroupCorrelationPlot(emb020, assay = "Spatial", cor = "nCount_Spatial_cor") + ggtitle("Log Normalization") +
  theme(plot.title = element_text(hjust = 0.5))
p2 <- GroupCorrelationPlot(emb020, assay = "SCT", cor = "nCount_Spatial_cor") + ggtitle("SCTransform Normalization") +
  theme(plot.title = element_text(hjust = 0.5))
p1 + p2

# Apply the function to each Seurat object
for (seurat_obj in seurat_objects) {
    # Process and plot for each object
    plot_output <- Normalize_Vs_SCTranform(seurat_obj)
    print(plot_output)
}
```

I think this shows that it is ok to continue with just the Seurat normalization.

# Run the default Seurat workflow
Running the default seurat workflow to just show the UMAPS for our data.

```{r IN PROGRESS (there is a bug), include=FALSE}

# Function for default Seurat processing
DefaultSeuratProcessing <- function(seurat_obj) {
    # Set default assay
    DefaultAssay(seurat_obj) <- "Spatial"

    # Apply Seurat processing steps
    seurat_obj <- NormalizeData(object = seurat_obj)
    seurat_obj <- FindVariableFeatures(object = seurat_obj)
    seurat_obj <- ScaleData(object = seurat_obj)
    seurat_obj <- RunPCA(object = seurat_obj)
    seurat_obj <- FindNeighbors(object = seurat_obj, dims = 1:30)
    seurat_obj <- FindClusters(object = seurat_obj)
    seurat_obj <- RunUMAP(object = seurat_obj, dims = 1:30)

    # Return the processed Seurat object
    return(seurat_obj)
}

# Apply the function to your Seurat objects
emb019 <- DefaultSeuratProcessing(emb019)
emb020 <- DefaultSeuratProcessing(emb020)
emb021 <- DefaultSeuratProcessing(emb021)
emb022 <- DefaultSeuratProcessing(emb022)
emb029 <- DefaultSeuratProcessing(emb029)
emb031 <- DefaultSeuratProcessing(emb031)
emb032 <- DefaultSeuratProcessing(emb032)

```
# Defining the gene set lists 

I used the gene set list below:

gene_sets <- list("DDX4", "FOXL2", "NR2F2", "WNT6", "LHX9", "EMX2", "NR5A1", "NR0B1", "AMHR2", "GREB1", "GATM", "IFI6", "CYP17A1", "GATA2", "RUNX1", "PAX8")

```{r}

# Define your gene sets
gene_sets <- list("DDX4", "FOXL2", "NR2F2", "WNT6", "LHX9", "EMX2", "NR5A1", "NR0B1", "AMHR2", "GREB1", "GATM", "IFI6", "CYP17A1", "GATA2", "RUNX1", "PAX8")
```

# Loop for the generation of the figures. 

```{r}
# Define the directory to save the plots
output_dir <- "results/MixedProbes/20240411/"

# Check if directory exists, if not create it
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# List of embryos to process
embryos <- list(emb019 = emb019, emb020 = emb020, emb021 = emb021, emb022 = emb022, emb029 = emb029, emb031 = emb031, emb032 = emb032)


# Loop through the embryos
for (emb_name in names(embryos)) {
  
  current_embryo <- embryos[[emb_name]]
  
  # Loop through the gene sets and plot each gene
  for (set_name in names(gene_sets)) {
    for (gene in gene_sets[[set_name]]) {
      
      # Use tryCatch to handle errors
      tryCatch({
        # Define the filename for the plot with the embryo name, gene set name, and gene name
        file_name <- paste0(output_dir, emb_name, "_", set_name, "_", gene, ".png")
        
        # Start the png device for plotting with increased resolution
        png(filename = file_name, width = 1600, height = 1200, res = 300)
        
        print(paste("Plotting", gene, "from", set_name, "set for embryo", emb_name))
        
        # Updated main formula
        plot <- SpatialFeaturePlot(current_embryo, features = gene, alpha = c(.1,5), pt.size.factor = 1) +
          theme(legend.position = "right") +
          scale_fill_viridis(option = "viridis")
        
        print(plot)  # This will print the plot to the graphics device
        
        # Close the png device, saving the plot to file
        dev.off()
        
      },
      error = function(e) {
        # Print the error message for this particular gene
        print(paste("Error plotting", gene, "from", set_name, "set for embryo", emb_name, ":", e$message))
      })
    }
  }
}

dev.off()
```

# Looking at a few examples
```{r}
p <- SpatialFeaturePlot(emb031, features = "FOXL2", alpha = c(.05,3), pt.size.factor = 1)
print(p)
```



```{r}

p <- SpatialFeaturePlot(emb020, features = c("AMHR2", "NR0B1", "STAR", "DPPA4", "LIN28A", "KISS1"), alpha = c(.05,3), pt.size.factor = 1) &
  theme(legend.position = "right") &
  scale_fill_viridis(option = "viridis")
print(p)

p <- SpatialFeaturePlot(emb021, features = c("AMHR2", "NR0B1", "STAR", "DPPA4", "LIN28A", "KISS1"), alpha = c(.05,3), pt.size.factor = 1) &
  theme(legend.position = "right") &
  scale_fill_viridis(option = "viridis")
print(p)
```


```{r}
sessionInfo()
```

